Mapping mutational effects along the evolutionary landscape of HIV envelope

### 2018-02-17 paper/ResponseToReviews.tex

{\color{black}
This is now fixed.
Previously we were running the analysis with a developmental version of \texttt{dms\_tools2} (version 2.2.dev1) that was not yet on \texttt{PyPI}.
Now the analysis uses a stable version (2.2.4) that is on \texttt{PyPI} and installable with \texttt{pip}.
Therefore, the analysis should now work if \texttt{dms\_tools2} is installed with \texttt{rplot} as described here: \url{https://jbloomlab.github.io/dms_tools2/installation.html#installing-the-rplot-module}.
Note that some other external Python packages are also needed as indicated by the imports at the beginning of the iPython notebook.
}

7. It might be helpful to increase the transparency on the points in Figure 3B, because at the current level it's impossible to resolve any detail in the blob near zero.

{\color{black}
We have added a new figure supplement (Figure 3--Figure supplement 2) that shows the correlations as a KDE contour plot rather than a scatter plot.
As is apparent from this new supplement, the vast majority of points are near the origin (low preference in both replicates). 
This is expected since most mutations are highly deleterious.
Therefore, the correlation signal is due to relatively small fraction of amino acids that have non-negligible preference---a fact that makes sesne, since most mutations are so deleterious as to be evolutionarily irrelevant.
Unfortunately, given the skew of the data it is not possible to find a transparency that strikes a good balance between showing the high-preference points clearly and not over-plotting the low preference points.
Therefore, we have elected to keep the original Figure 3B (scatter plot) while also adding the new contour plot in Figure 3--Figure supplement 2) to show both elements of the data.
}

8. In Figure 4, it's not clear why the p-value for $\omega$ greater/less than one is used for shading rather than the estimated $\omega$ itself.

{\color{black}
We have added text in the legend of Figure 4 to explain this.
Briefly, for individual sites, the point estimates of dN/dS values (or $\omega_r$) are not reliable when using a FEL framework for the reasons described in Pond et al (2015) and Murrell et al (2012).

Here is the basic intuition. 
Imagine two sites: one has 3 nonsynonymous substitutions and 0 synonymous substitutions, while the other has 20 nonsynonymous substitutions and 2 synonymous substitutons.
If we just calculated $\omega_r$, we would get a value of infinity for the first site and 10 for the second site.
But in reality, we think there is substantially stronger evidence for diversifying selection at the second site, since there are many more nonsynonymous substitutions.
In an FEL framework, we assess the statistical evidence that $\omega_r \ne 1$, and find much more evidence for the second site.
This increased evidence is manifested by the larger $P$-value rather than the value of $\omega_r$ itself.}

### 2018-02-17 paper/manuscript.tex
explain why we report P rather than omega_r

\begin{figure}
\begin{fullwidth}
{\includegraphics[width=1.3\textwidth]{figures/BG505_prefs.pdf}}
\caption{\label{fig:BG505prefs}
Amino-acid preferences for the BG505 Env.
At each site, the height of the letter is proportional for that site's preference for that amino acid.
The top color bar indicates the region of Env (gp120 variable loop, gp120 not variable loop, or gp41).
The lower color bar indicates the evidence that the site evolves faster ($\omega_r > 1$) or slower ($\omega_r < 1$) than expected given the experiments~\citep[see][]{bloom2017identification}.
The letters above the logos indicate the wildtype amino acid in BG505.
Sites are numbered using the HXB2 scheme~\citep{korber1998numbering}).
This logo plot shows the site-specific amino-acid preferences for BG505 after averaging the replicates and re-scaling by the stringency parameter in \TABLE{phydms}.
The figure was generated using \texttt{dms\_tools2}~\citep[\url{https://jbloomlab.github.io/dms_tools2/}]{bloom2015software}, which in turn utilizes \texttt{weblogo}~\citep{crooks2004weblogo}.
}
\figdata{The numerical values of the amino-acid preferences plotted in this figure are in \texttt{rescaled\_BG505\_prefs.csv.}}
\figdata{The sequence of BG505 Env and mapping from sequential (\emph{original} column) to HXB2 numbering (\emph{new} column) is in \texttt{BG505\_to\_HXB2.csv}.}
\figdata{The $\omega_r$ values and associated $P$-values for BG505 in HXB2 numbering are in \texttt{BG505\_omegabysite.tsv}.}
\end{fullwidth}
\end{figure}

>>>>

\begin{figure}
\begin{fullwidth}
{\includegraphics[width=1.3\textwidth]{figures/BG505_prefs.pdf}}
\caption{\label{fig:BG505prefs}
Amino-acid preferences for the BG505 Env.
At each site, the height of the letter is proportional for that site's preference for that amino acid.
The top color bar indicates the region of Env (gp120 variable loop, gp120 not variable loop, or gp41).
The lower color bar indicates the evidence that the site evolves faster ($\omega_r > 1$) or slower ($\omega_r < 1$) than expected given the experiments~\citep[see][]{bloom2017identification}; we report the $P$-value for $\omega_r \ne 1$ rather than the value of $\omega_r$ itself since point estimates of these values are unreliable for individual sites due to low numbers of observations~\citep{kosakovsky2005not,murrell2012detecting}.
The letters above the logos indicate the wildtype amino acid in BG505.
Sites are numbered using the HXB2 scheme~\citep{korber1998numbering}).
This logo plot shows the site-specific amino-acid preferences for BG505 after averaging the replicates and re-scaling by the stringency parameter in \TABLE{phydms}.
The figure was generated using \texttt{dms\_tools2}~\citep[\url{https://jbloomlab.github.io/dms_tools2/}]{bloom2015software}, which in turn utilizes \texttt{weblogo}~\citep{crooks2004weblogo}.
}
\figdata{The numerical values of the amino-acid preferences plotted in this figure are in \texttt{rescaled\_BG505\_prefs.csv.}}
\figdata{The sequence of BG505 Env and mapping from sequential (\emph{original} column) to HXB2 numbering (\emph{new} column) is in \texttt{BG505\_to\_HXB2.csv}.}
\figdata{The $\omega_r$ values and associated $P$-values for BG505 in HXB2 numbering are in \texttt{BG505\_omegabysite.tsv}.}
\end{fullwidth}
\end{figure}
